\documentclass[11pt]{article}

% # FONTS

\usepackage[T1]{fontenc}
\renewcommand*\ttdefault{lmvtt}
\usepackage[bitstream-charter]{mathdesign}
\usepackage{amsmath}
\usepackage[scaled=0.92]{PTSans}

% # GEOMETRY

\usepackage[
  paper  = letterpaper,
  left   = 1.0in,
  right  = 1.0in,
  top    = 1.25in,
  bottom = 1.25in,
]{geometry}
\usepackage[final,expansion=alltext]{microtype}
\usepackage[english]{babel}
\usepackage[parfill]{parskip}
\usepackage{afterpage}
\usepackage{framed}
\usepackage{enumitem}
\usepackage{pifont}     % needed for dingbats for colophon

% # COLOR

\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{shadecolor}{gray}{0.9}

% # BIBLIOGRAPHY

\usepackage{bibentry}
\makeatletter\let\saved@bibitem\@bibitem\makeatother
\let\oldbibentry\bibentry
\renewcommand{\bibentry}[1]{\oldbibentry{#1}.}
\def\FormatName#1{%
  \def\myname{D. J. Saunders}%
  \edef\name{#1}%
  \ifx\name\myname
    \textbf{#1}%
  \else
    #1%
  \fi
}

\usepackage{hyperref}
\makeatletter\let\@bibitem\saved@bibitem\makeatother

% # HEADERS

\usepackage{fancyhdr}
\fancyhf{}
\pagestyle{fancy}
\lhead{\textcolor{gray}{\emph{Daniel J. Saunders}}}
\rhead{\textcolor{gray}{\thepage}}
\renewcommand{\headrulewidth}{0pt}

% Renew section
\renewcommand{\section}[1]
{\vspace{1.0\baselineskip}{\Large\textbf{#1}}}

% Generate colophon at the end of the document
\newcommand{\makecolophon}
{\centering\ding{118} Last modified on \today~\ding{118}}

\begin{document}
\bibliographystyle{plain-custom}
\nobibliography{bib}
\thispagestyle{empty}

\parbox[c][0.5in]{3.5in}{\huge \textbf{Daniel J. Saunders}}
\parbox[c][0.5in]{2.5in}{%
\small
\vspace{20ex}
\texttt{djsaunde@cs.umass.edu}
\texttt{https://djsaunde.github.io}
\texttt{https://github.com/djsaunde}
\texttt{https://medium.com/@danjsaund/}
}

M.Sc. Student \\
BINDS Lab \\
University of Massachusetts, Amherst \\
College of Information and Computer Sciences \\
Amherst, MA

\section{Education}

\begin{itemize}[leftmargin=12pt]
  \item[] M.Sc. Computer Science, University of Massachusetts, Amherst \hfill 2017-- (Expected 12/2018)\\
  Concentration: machine learning, computational neuroscience \\
  GPA: 3.7
  \item[] B.S. Computer Science, University of Massachusetts, Amherst \hfill 2012--2016 \\
  Concentration: theoretical computer science, artificial intelligence \\
  GPA: 3.5
  \item[] B.S. Mathematics, University of Massachusetts, Amherst \hfill 2012--2016 \\
  Concentration: mathematical computing \\
  GPA: 3.5
\end{itemize}

\section{Employment}

\begin{itemize}[leftmargin=12pt]
  \item[] Graduate Research Assistant \hfill Spring 2017--Fall 2018 \\
  \textit{Biologically Inspired Neural and Dynamical Systems Lab} \\
  Supervisor(s): Professors Robert Kozma and Hava Siegelmann
  \item[] Data Science Intern \hfill Summer 2018 \\
  \textit{HealthcareSource} \\
  Supervisor(s): Patrick McDonough
  \item[] Programmer \hfill Fall 2016--Fall 2018 \\
  \textit{Department of Resource Economics, University of Massachusetts, Amherst} \\
  Supervisor(s): Professors Christian Rojas and Debi Mohapatra
  \item[] Research Intern \hfill Summer 2017 \\
  \textit{Air Force Research Lab Automatic Target Recognition Center} \\
  Supervisor(s): Dr. Roman Ilin and Professor Robert Kozma
  \item[] Programmer \hfill Summer 2016 \\
  \textit{Biologically Inspired Neural and Dynamical Systems Lab} \\
  Supervisor(s): Dr. Hava Siegelmann
  \item[] Programmer \hfill Summer 2015--Winter 2017 \\
  \textit{Cognition and Action Lab, University of Massachusetts, Amherst} \\
  Supervisor(s): Professor Rebecca Spencer
  \item[] Software Development Intern \hfill Summer 2015 \\
  \textit{Epsilon} \\
  Supervisor(s): Patrick McDonough
\end{itemize}

\section{Awards}

\begin{itemize}[leftmargin=12pt]
  \item[] Bay State Master's Program (50\% of tuition \& fees) \hfill Spring 2017- Fall 2018
  \item[] Graduate Research Fellowship \hfill Spring 2017--Fall 2018
  \item[] IJCNN 2018 - Runner-Up for Best Paper Award \hfill Spring 2018 \\ \textit{Unsupervised Learning with Self-Organizing Spiking Neural Networks}
\end{itemize}

\section{Publications}

\textsc{Journal Articles}
\begin{enumerate}
  \item \bibentry{10.3389/fninf.2018.00089}
\end{enumerate}

% \textsc{Books}
% \textsc{Edited Volumes}
% \textsc{Book Chapters}

\textsc{Conference Articles}
\begin{enumerate}[resume]
  \item \bibentry{hazan2018}
  \item \bibentry{saunders2018}
\end{enumerate}

\textsc{Working papers (working titles)}
\begin{enumerate}[resume]
	\item H. Hazan, D. J. Saunders, D. T. Sanghavi, H. T. Siegelmann, and R. Kozma. Lattice map spiking neural networks for clustering and classifying image data. Submitted to \textit{Annals of Mathematics and Artificial Intelligence}.
	\item D. J. Saunders, H. Hazan, H. T. Siegelmann, and R. Kozma. Locally Connected Spiking Neural Networks for Unsupervised Feature Learning.
	\item H. Hazan, D. Patel, D. J. Saunders, H. T. Siegelmann, and R. Kozma. Transferring reinforcement learning policies to spiking neural networks for Atari game-playing.
\end{enumerate}

% \textsc{Workshop Articles}
% \textsc{Acknowledgements}

\section{Talks \& Presentations}

\textsc{Poster presentations}
\begin{enumerate}
	\item D. J. Saunders, H. Hazan, H. Khan, H. T. Siegelmann, R. Kozma. \href{https://djsaunde.github.io/files/BindsNET_PyTorch_conference_poster.pdf}{BindsNET: An ML-oriented spiking networks library built with PyTorch}. PyTorch Developers Conference 2018. San Francisco, California. October 2, 2018.
\end{enumerate}

\section{Software}

\begin{enumerate}
  \item BindsNET: A spiking neural networks simulation library built with PyTorch. \\
  GitHub repo: \href{https://github.com/Hananel-Hazan/bindsnet}{https://github.com/Hananel-Hazan/bindsnet} \\
  Authors: \textbf{D. J. Saunders}, H.~Hazan, and H.~Khan.
  
  \item NYC-TLC: A Python package for the downloading and manipulation of NYC taxi trip records using the Dask distributed computation library. \\
  GitHub repo: \href{https://github.com/djsaunde/nyctlc}{https://github.com/djsaunde/nyctlc} \\
  Authors: \textbf{D. J. Saunders}
\end{enumerate}

\section{Technical Skills, etc.}

\begin{itemize}
  \item Programming languages (ordered by decreasing proficiency): Python, R, Java, C/C++, MATLAB, SQL, JavaScript, Haskell
  \item Machine learning frameworks (ordered by decreasing proficiency): PyTorch, Scikit-Learn, Keras, Tensorflow, Theano, MatConvNet
  \item Reading list: \href{https://djsaunde.github.io/read/read.pdf}{https://djsaunde.github.io/read/read.pdf}
  \item Notes on lab work, books, and papers: \href{https://djsaunde.github.io/notes}{https://djsaunde.github.io/notes}
\end{itemize}

\section{Relevant coursework}

\begin{itemize}
  \item Machine learning (undergraduate \& graduate)
  \item Artificial intelligence (undergraduate \& graduate)
  \item Deep neural networks
  \item Statistics I \& II (undergraduate)
  \item Mathematical statistics I \& II (graduate)
  \item Distributed \& operating systems
  \item Causal inference
  \item Algorithms for data science
  \item Applied information theory
  \item Dynamical systems
  \item Computational complexity
  
\end{itemize}

% \section{Popular Press}
% \section{Awarded Grants}
% \section{Industry}
% \section{Teaching}
% \section{Professional Service}

% \textsc{Senior Program Committee}
% \textsc{Associate Editor and Editorial Board}
% \textsc{Journal Reviewing}
% \textsc{Conference Reviewing}
% \textsc{Workshop Organization}
% \textsc{Professional Memberships}
% \textsc{Grant Reviewing}
% \textsc{Former Ph.D. Students}
% \textsc{Current Ph.D. Students}
% \textsc{Postdoctoral Fellows}
% \textsc{Other Ph.D. Thesis Committees}
% \textsc{Mentoring}
% \textsc{Advising and Consulting}

% \section{Invited Talks and Panels}

% \vfill
% \makecolophon

\end{document}
